---
title: "Next generation method"
author: "Damie Pak"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The mathematical formulation
For calculating the reproductive number ($R0$) or the effective reproductive number, we can use the next generation method (NGM). Here, we create an *$F$* matrix where we track how new infections in class $i$ emerge. We also create a $*V*$ matrix where we record how individuals leave the infectious stage. By calculating the eigenvalue of ($FV^-1$), we should be able to calculate the $R0$ by calculating the NGM at the disease-free equilibirum. 

This is fairly simple doing it in a single-patch system, but becomes a bit harder when we have multiple patches and multiple vectors. Here, we show step-by-step on how to calculate the R-effective of the individual patches. 

## One-host and one vector on a 2-patch system

Assumptions

1. There are two patches and assume we do not care about the distance

2. The host human population do not disperse and is kept constant over time 

3. The vector population (both susceptible and infected) disperses between PATCH 1
and PATCH 2

The full model can be written as below:

$$
\frac{dH_{S(j)}}{dt} = \wedge N_j - \lambda_V H_S \frac{V_I}{N_V} - \mu_H H_{S(j)}
$$

$$
\frac{dH_{I(j)}}{dt} = \lambda_V H_S \frac{V_I}{N_V} - \gamma H_{S(j)} - \mu_H H_{S(j)}
$$

$$
\frac{dV_{S(j)}}{dt} = BN_V - \lambda_H V_S \frac{H_I}{N_H} - \mu_V V_{S(j)} - a V_{S(j)} +aP_{S(k)}
$$ 

$$
\frac{dV_{I(j)}}{dt} = \lambda_H V_S \frac{H_I}{N_H} - \mu_V V_{I(j)} - a V_{I(j)} +
a V_{I(k)}
$$

To construct the NGM of the entire metapopulation. Let us think of this way. We have two patches and 2 classes of infected ($H_I$ and $V_I$). If $n$ represents the number of patches and $c$ represents the infected classes, we should have a $n \times c$ matrix.

Specifically,

$$
\begin{bmatrix}
H_{I(1)} \leftarrow H_{I(1)} & H_{I(1)} \leftarrow H_{I(2)}  & H_{I(1)} \leftarrow 
V_{I(1)}  &  H_{I(1)} \leftarrow 
V_{I(2)}\\
H_{I(2)} \leftarrow H_{I(1)} & H_{I(2)} \leftarrow H_{I(2)}  & H_{I(2)} \leftarrow 
V_{I(1)}  &  H_{I(2)} \leftarrow 
V_{I(2)}\\
V_{I(1)} \leftarrow H_{I(1)} & V_{I(1)} \leftarrow H_{I(2)}  & V_{I(1)} \leftarrow 
V_{I(1)}  &  V_{I(1)} \leftarrow 
V_{I(2)}\\
V_{I(2)} \leftarrow H_{I(1)} & V_{I(2)} \leftarrow H_{I(2)}  & V_{I(2)} \leftarrow 
V_{I(1)}  &  V_{I(2)} \leftarrow 
V_{I(2)}\\
\end{bmatrix}
$$

Because infected humans and vectors do not directly infect other humans and vectors in the same- the
diagonals can all be 0. Additionally, because infectious humans do not disperse, we can make these entries 0.

$$
\begin{bmatrix}
0&0  & H_{I(1)} \leftarrow 
V_{I(1)}  &  0\\
0 & 0  &0  &  H_{I(2)} \leftarrow 
V_{I(2)}\\
V_{I(1)} \leftarrow H_{I(1)} & 0 & 0 &  V_{I(1)} \leftarrow 
V_{I(2)}\\
0& V_{I(2)} \leftarrow H_{I(2)}  & V_{I(2)} \leftarrow 
V_{I(1)}  & 0\\
\end{bmatrix}
$$

For $H_{I(j)} \leftarrow V_{I(j)}$, we have $\lambda_V \frac{H_{S(j)}}{N_{V(j)}}$.For $V_{I(j)} \leftarrow H_{I(j)}$, we have $\lambda_H \frac{V_{S(j)}}{N_{H(j)}}$

$$
\textbf{F} =\begin{bmatrix}
0&0  &\lambda_V \frac{H_{S(1)}}{N_{V(1)}} &  0\\
0 & 0  &0  &  \lambda_V \frac{H_{S(2)}}{N_{V(2)}}\\
\lambda_H \frac{V_{S(1)}}{N_{H(1)}} & 0 & 0 & a\\
0& \lambda_H \frac{V_{S(2)}}{N_{H(2)}} & a & 0\\
\end{bmatrix}
$$

The V matrix is: 
$$
\textbf{V} =\begin{bmatrix}
-(\gamma + \mu_H)  & 0 &  0 &0\\
0 &  -(\gamma + \mu_H)  &0  &0 \\
0 & 0 & -(a + \mu_V) & 0\\
0& 0 & 0 & -(a + \mu_V)\\
\end{bmatrix}
$$

Therefore the $V^{-1}$ matrix is:

$$
\begin{bmatrix}
-\frac{1}{\gamma + \mu_H} & 0 & 0 & 0 \\
0 & -\frac{1}{\gamma + \mu_H} & 0 & 0 \\
0 & 0 & -\frac{1}{a + \mu_V} & 0 \\
0 & 0 & 0 & -\frac{1}{a + \mu_V}
\end{bmatrix}
$$


the matrix
$$
{FV^{-1}} =\begin{bmatrix}
0&0  &\frac{\lambda_V  H_{S(1)}}{N_{V(1)} (a+ \mu_V)} &  0\\
0 & 0  &0  &  \frac{\lambda_V H_{S(2)}}{N_{V(2)}(a+ \mu_V)}\\
 \frac{\lambda_H V_{S(1)}}{N_{H(1)}(\gamma + \mu_H)} & 0 & 0 & \frac{a}{a+\mu_V}\\
0&  \frac{\lambda_H V_{S(2)}}{N_{H(2)}(\gamma + \mu_H)} & \frac{a}{a + \mu_V} & 0\\
\end{bmatrix}
$$
It is most likely that the equation is going to be horrible solving by hand so let's 
code this up:

## The desolve code

Let's write the first system of equations in a desolve code 

```{r}
metapopulation_model <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
          
        b_H <- matrix(rep_len(parameters["b_H"], 2), ncol = 1) # Human birth rate
        b_V<- matrix(rep_len(parameters["b_V"], 2), ncol = 1)  #P.vector birth rate
        mu_H <- matrix(rep_len(parameters["mu_H"], 2), ncol = 1)  #
        mu_V <-  matrix(rep_len(parameters["mu_V"], 2), ncol = 1)  #P. vector death rate
          
        lambda_H <- matrix(rep_len(parameters["lambda_H"], 2), ncol = 1)
        lambda_V <- matrix(rep_len(parameters["lambda_V"], 2), ncol = 1) 
        
        a <- matrix(c(0,0,1,0),ncol =2,byrow = TRUE)
       ###The metapopulation
        H_S <- matrix(state[1:2], ncol = 1) #Susceptible human
        H_I <- matrix(state[3:4], ncol = 1) #Infected human
        H_R <- matrix(state[5:6], ncol = 1) #Recovered human
        V_S <- matrix(state[7:8], ncol = 1) #Susceptible 
        V_I <- matrix(state[9:10], ncol = 1) #Infected primary
  
        N_H <- H_S + H_I + H_R
        N_V <- V_S + V_I
            # Differential equations
    
        dH_S <-  (b_H * N_H) - (lambda_V * H_S * (V_I/N_V)) - (mu_H * H_S)     
        
        dH_I <- (lambda_V * H_S * (V_I/N_V)) - (gamma  + mu_H) * H_I  
        
        dH_R <-  (gamma  - mu_H) * H_I  

        dV_S <-  (b_V * N_V) - (lambda_H * V_S * (H_I/N_H)) - (mu_H * V_S)  + 
                (a %*% V_S) - (colSums(a) * V_S)
        
        dV_I <- (lambda_H * V_S * (H_I/N_H))  - (mu_H * V_I)  +
                (a %*% V_I) - (colSums(a) * V_I)

    # Return the rates of change
    return(list(c(dH_S, dH_I, dH_R, dV_S, dV_I)))
  })
}

```

```{r}
initial_y <- c(HS =  rep(100, 2),
               HI = rep(0,2),
               HR = rep(0,2),
               VS = rep(100,2),
               VI =rep(10,2))

 parameters_full <-  c(b_H = 0.05, #Human birth rate
                b_V = (0.05), #P.vector birth rate
                mu_H = 0.01, #Human death rate
                mu_V = 0.05, #P. vector death rate

                lambda_V= 0.132, #transmission probability of p. vector
                lambda_H = 0.132 * 0.55, #transmission probability of s. vector

                # Recovery rate of the acute phase
                gamma = 1/56)
```                


```{r}

deSolve_results <-data.frame(ode(
                times = seq(1,100,1),
                y =   initial_y  ,
                func = metapopulation_model,
                parms = parameters_full,
                method = 'lsodar'))
```

## Calculating the R effective

```{r}
HS_1 <- deSolve_results[,"HS1"]
HS_2 <- deSolve_results[,"HS2"]

HI_1 <- deSolve_results[,"HI1"]
HI_2 <- deSolve_results[,"HI2"]

HR_1 <- deSolve_results[,"HR1"]
HR_2 <- deSolve_results[,"HR2"]

NH_1 <- HS_1 + HI_1 + HR_1
NH_2 <- HS_2 + HI_2 + HR_2


VS_1 <- deSolve_results[,"VS1"]
VS_2 <- deSolve_results[,"VS2"]

VI_1 <- deSolve_results[,"VI1"]
VI_2 <- deSolve_results[,"VI2"]

NV_1 <-  VS_1 + VI_1 
NV_2 <- VS_2 + VI_2

```

```{r}
R_effective_calculator_list<- NULL
New_addition_list <- NULL
for (i in seq(1,100)){

HI1_from_VI1 <-  (parameters_full[["lambda_V"]] * HS_1[i])/ 
        (NV_1[i] * (0 + parameters_full[["mu_V"]]))
HI2_from_VI2 <-  (parameters_full[["lambda_V"]] * HS_2[i])/ 
        (NV_2[i] * (1 + parameters_full[["mu_V"]]))

VI1_from_HI1 <- (parameters_full[["lambda_H"]] * VS_1[i])/ 
        (NH_1[i] * (parameters_full[["gamma"]] + parameters_full[["mu_H"]]))

VI1_from_VI2 <- 1/(1+ parameters_full[["mu_V"]])
VI2_from_VI1<- 0/(1+ parameters_full[["mu_V"]])

VI2_from_HI2 <- (parameters_full[["lambda_H"]] * VS_2[i])/ 
        (NH_2[i] * (parameters_full[["gamma"]] + parameters_full[["mu_H"]]))

FV_1 <- matrix(c(0,0,HI1_from_VI1,0,
         0,0,0,HI2_from_VI2,
       VI1_from_HI1 ,0,0,VI1_from_VI2,
       0,VI2_from_HI2,VI2_from_VI1,0),ncol = 4)


R_effective_calculator_list[[i]] <- max(eigen(FV_1)$values)
New_addition_list [[i]]<-cbind.data.frame(states = c("HI1","HI2","VI1","VI2"), new =rowSums(FV_1 ),time = i)

}

```


```{r}
R_effective_calculator_df <- cbind.data.frame(time=seq(1,100,1),
                                   R_eff=unlist(R_effective_calculator_list))
```

```{r}
plot(R_effective_calculator_df$time,
     R_effective_calculator_df$R_eff)

```


```{r}
New_addition_df <- do.call(rbind, New_addition_list)
```


```{r}
ggplot(New_addition_df, aes( x= time, y =log10(new) , color = states))+
        geom_line()
```
