---
title: "Simulating the network in tritonet"
author: "Damie Pak"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r include=FALSE}
library(ggplot2)
library(ggnetwork)
library(igraph)
library(reshape2)
```

This is a step by step guide for how we simulate the spatial network for 
`tritonet`.

## 1. Sample the x and y coordinates for the patches.

```{r}
set.seed(24601) # Set the seed number
max_distance <- 20 # Set the maximum limit of the xy plane.
```  
    
We use the `sample` function to randomly select both the x (longitude) and y (latitude) 
coordinates for each node. Using `dist`, we can then calculate the distance
matrix for the pairwise distances between all nodes. The weight of the
edges is then calculated using a negative exponential kernel. 

```{r}
xy <- seq(1, max_distance, length.out = 2000) ### All possible coordinates
x_coord <- sample(xy, 100, replace = TRUE) # x-coordinate
y_coord <- sample(xy, 100, replace = TRUE) # y-coordinate
xy_coord <- cbind(x_coord, y_coord) # xy-coordinates combined
NegExpDist <- as.matrix(exp(-dist(xy_coord))) # distance matrix with neg. exp kernel
```

## 2. Convert the distance matrix into an adjacency matrix

```{r}
Adj_graph <- graph_from_adjacency_matrix(NegExpDist,
  mode = "undirected",
  diag = FALSE,
  weighted = TRUE
)

## Adding latitude and longitude
V(Adj_graph)$Long <- xy_coord[, 1] # x-coordinates
V(Adj_graph)$Lat <- xy_coord[, 2] # y-coordinates
```

When we plot the network, we can see that all nodes are connected to each other.

```{r echo=FALSE, fig.height=10, fig.width=10}
ggplot(ggnetwork(Adj_graph), aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(alpha = 0.1) +
  geom_nodes(color = "red", size = 10) +
  geom_nodetext(aes(label = name),
    fontface = "bold", color = "white"
  ) + theme_void()
```

## 3. Reduce the edges

We are going to delete the majority of the edges. We assume a very low connectance (1%) and
we back-calculate the number of edges that we must keep.

```{r} 
number_of_edges <- (0.01 * (100^2))
```   

We choose the top 100 highest-weight edges and delete all other edges.

```{r}
### If the number of edges required for connectance is 100, then
### choose the 100 likeliest (highest weight) edges.
deleted_edges_graph <- delete_edges(
  Adj_graph,
  which(E(Adj_graph)$weight < sort(E(Adj_graph)$weight,
    decreasing = T
  )[number_of_edges])
)
```

This is what it looks like now; we can see that there are components

```{r echo=FALSE}
ggplot(ggnetwork(deleted_edges_graph), aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(alpha = 0.7) +
  geom_nodes(color = "red", size = 6) +
  geom_nodetext(aes(label = name),
    fontface = "bold", color = "white"
  ) +
  theme_void()
```

## 4. Choose the component with the greatest number of nodes

Using the function `decompose`, we can split the network into smaller networks.
```{r}
decomposed_components <- decompose(deleted_edges_graph)

# Count the number of nodes for each component and then give me
### the index for the largest component.

biggest_component_length <- which.max(lapply(
  decomposed_components,
  function(x) {
    vcount(x)
  }
))

### retrieve our network of interest
network_of_interest <- decomposed_components[[biggest_component_length]]
```

This is the network of interest (the biggest component)

```{r echo=FALSE}
gg_network_1 <- ggplot(ggnetwork(network_of_interest), aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(alpha = 0.7) +
  geom_nodes(color = "red", size = 6) +
  geom_nodetext(aes(label = name),
    fontface = "bold", color = "white"
  ) +
  theme_void()
gg_network_1
```

## 5. Calculate the new connectance

I create a function called `connectance_calculator` that calculates the connectance 
when given the number of nodes and edges.
```{r}
### calculate the connectance by inputting the number of nodes and the number of
### edges

connectance_calculator <- function(nodes, edges) {
  return(edges / (nodes^2))
}
```


```{r}
connectance_calculator(
  vcount(network_of_interest),
  ecount(network_of_interest)
)
```

## 6. Create an empty list to populate with igraph objects

Let's create a list:

```{r}
adj_list <- NULL # For the actual igraph
adj_info_list <- NULL # For information about each igraph
```

Let's manually add the first one in.

```{r}
adj_list[[1]] <- network_of_interest #put the igraph object in.


adj_info_list[[1]] <- c( 
  num_nodes = vcount(network_of_interest),
  num_edges = ecount(network_of_interest),
  connectance = connectance_calculator(
    vcount(network_of_interest),
    ecount(network_of_interest)
  )
)
```

## 7. Add edges one by one

First, we calculate the new distance matrix of the component network of interest:

```{r}
# Get the x-y coordinates
xy_coord_interest <- cbind(
  V(network_of_interest)$Long,
  V(network_of_interest)$Lat
)

# Calculate new distance matrix
DispMat_interest <- as.matrix(exp(-dist(xy_coord_interest)))
```

We get the edge list of the network of interest.
```{r}
edgelist_of_interest <- as_edgelist(network_of_interest, names = F)

### The columns show the patches that are connected by an edge
head(edgelist_of_interest)
```

By melting the distance matrix, we can then get a data.frame that shows the 
edge connections between the different nodes as well as the edge weights.

```{r}
melted_edge_list <- melt(DispMat_interest)

### patch1, patch2, and weight are the new column names
colnames(melted_edge_list) <- c("patch1", "patch2", "weight")
```

We want to remove rows from `melted_edge_list` that already exist
in the `network_of_interest`. This is because we're interested in adding new edges that
do not currently exist in the network.


```{r}
new_distance <- subset(
  melted_edge_list,
  !(paste0(
    melted_edge_list$patch1, "-",
    melted_edge_list$patch2
  )
  %in%
    paste0(
      edgelist_of_interest[, 1],
      "-", edgelist_of_interest[, 2]
    )
  )
)
```

Let us order the new data.frame by the edge weight:

```{r}
new_distance <- new_distance[order(new_distance$weight, decreasing = TRUE), ]
head(new_distance)
```

We're going to loop this, but just to how what is happening. We add an edge 
between patch1 and patch2 as well as its weight.

```{r}
network_of_interest_added <- network_of_interest + edge(
  c(
    new_distance[1, "patch1"],
    new_distance[1, "patch2"]
  ),
  weight = new_distance[1, "weight"]
)
```

Again, we're going to automate this, but we are going to add the information we need
to the lists that we made earlier.

```{r}
adj_list[[2]] <- network_of_interest_added
adj_info_list[[2]] <- c(
  num_nodes = vcount(network_of_interest_added),
  num_edges = ecount(network_of_interest_added),
  connectance = connectance_calculator(
    vcount(network_of_interest_added),
    ecount(network_of_interest_added)
  )
)
```

## 8. Loop through.

```{r}
for (i in seq(2, nrow(new_distance))) {
  network_of_interest_added <- network_of_interest_added + edge(
    c(
      new_distance[i, "patch1"],
      new_distance[i, "patch2"]
    ),
    weight = new_distance[i, "weight"]
  )


  adj_list[[i + 1]] <- network_of_interest_added
  adj_info_list[[i + 1]] <- c(
    num_nodes = vcount(network_of_interest_added),
    num_edges = ecount(network_of_interest_added),
    connectance = connectance_calculator(
      vcount(network_of_interest_added),
      ecount(network_of_interest_added)
    )
  )
}
```

## 9. Check that there is a positive relationship with edge number and connectance

```{r}
adj_info_df <- data.frame(do.call(rbind, adj_info_list))
```


```{r echo=TRUE, fig.height=3, fig.width=5}
ggplot(adj_info_df, aes(x = num_edges, y = connectance)) +
  geom_point() +
  ylab("Connectance") +
  xlab("Number of edges") +
  theme_classic() +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 15)
  )
```

We can see that by increasing the number of edges, we also increase the connectance.

## 10. Creating the full function

Going to be a huge function so break it into much smaller sub-functions.

The function `simulate_xy_coordinates` corresponds to Step 1 (Sample x-y coordinates for the patches). The output should be a list with the first element being the data.frame holding the x and y coordinates of the nodes and the second element being the distance matrix.
```{r}
simulate_xy_coordinates <- function(seed = 24601, max_distance) {
  set.seed(seed)
  xy <- seq(1, max_distance, length.out = 2000) ### List of all possible coordinates
  x_coord <- sample(xy, 100, replace = TRUE) # x-coordinate
  y_coord <- sample(xy, 100, replace = TRUE) # y-coordinate
  xy_coord <- cbind(x_coord, y_coord) # xy-coordinates combined
  NegExpDist <- as.matrix(exp(-dist(xy_coord))) # distance matrice with kernel

  return(list(xy_coord, NegExpDist))
}
```

The function `retrieve_biggest_component` corresponds to Step 2 (Convert the distance matrix into an adjacency matrix), Step 3 (Reduce the edges), and Step 4 (Choose the components with the greatest number of nodes). The input takes the list element from `simulate_xy_coordinate` and returns the network of interest.

```{r}
retrieve_biggest_component <- function(list) {
  Adj_graph <- graph_from_adjacency_matrix(list[[2]],
    mode = "undirected",
    diag = FALSE,
    weighted = TRUE
  )

  ## Adding latitude and longitude
  V(Adj_graph)$Long <- list[[1]][, 1] # x-coordinates
  V(Adj_graph)$Lat <- list[[1]][, 2] # y-coordinates

  number_of_edges <- (0.01 * (100^2))

  deleted_edges_graph <- delete_edges(
    Adj_graph,
    which(E(Adj_graph)$weight < sort(E(Adj_graph)$weight,
      decreasing = T
    )[number_of_edges])
  )


  decomposed_components <- decompose(deleted_edges_graph)

  # Count the number of nodes for each componenent and then give me
  ### the index for the largest.
  biggest_component_length <- which.max(lapply(
    decomposed_components,
    function(x) {
      vcount(x)
    }
  ))

  ### retrieve our network of interest
  network_of_interest <- decomposed_components[[biggest_component_length]]

  return(network_of_interest)
}
```
The function `recalculate_distance_matrix` correspond to the first half of step 5 (Add edges one by one). You input the network of interest and should return a data.frame that has all the possible edges (that are not in the current network) sorted in decreasing order of edge weight.

```{r}
recalculate_distance_matrix <- function(network) {
  # Get the x-y coordinates
  xy_coord_interest <- cbind(
    V(network)$Long,
    V(network)$Lat
  )

  # Calculate new distance matrices
  DispMat_interest <- as.matrix(exp(-dist(xy_coord_interest)))

  edgelist_of_interest <- as_edgelist(network, names = F)

  melted_edge_list <- melt(DispMat_interest)

  colnames(melted_edge_list) <- c("patch1", "patch2", "weight")


  new_distance <- subset(
    melted_edge_list,
    !(paste0(
      melted_edge_list$patch1, "-",
      melted_edge_list$patch2
    )
    %in%
      paste0(
        edgelist_of_interest[, 1],
        "-", edgelist_of_interest[, 2]
      )
    )
  )

  new_distance_df <- new_distance[order(new_distance$weight, decreasing = TRUE), ]

  return(new_distance_df)
}
```

The full function thus looks like this:

```{r}
simulate_spatial_network <- function(seed, max_distance) {
        
  list_xy_coord <- simulate_xy_coordinates(seed, max_distance)
  network_interest <- retrieve_biggest_component(list_xy_coord)
  possible_edges_df <- recalculate_distance_matrix(network_interest)
  adj_list <- NULL
  adj_info_list <- NULL

  ### Manually add the first network in

  adj_list[[1]] <- network_interest
  adj_info_list[[1]] <- c(
    num_nodes = vcount(network_interest),
    num_edges = ecount(network_interest),
    connectance = connectance_calculator(
      vcount(network_interest),
      ecount(network_interest)
    )
  )

  ### For loop time
  for (new_edge in seq(1, nrow(possible_edges_df))) {
    network_interest <- network_interest + edge(c(new_distance[new_edge, "patch1"], new_distance[new_edge, "patch2"]),
      weight = new_distance[new_edge, "weight"]
    )


    adj_list[[new_edge + 1]] <- network_interest
    adj_info_list[[new_edge + 1]] <- c(
      num_nodes = vcount(network_interest),
      num_edges = ecount(network_interest),
      connectance = connectance_calculator(
        vcount(network_interest),
        ecount(network_interest)
      )
    )
  }
  return(list(adj_list, do.call(rbind, adj_info_list)))
}
```
   
## 11. Testing the full function                
```{r}
simulated_list <- simulate_spatial_network (24601, 20)
```

```{r}
print(simulated_list[[2]])
```
         
